### 2017-10-20 papers/main.tex

\documentclass{llncs}

\usepackage{makeidx}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{indentfirst}
\usepackage{verbatim}
\usepackage{amsmath, amssymb}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{url}
\usepackage{stmaryrd}
\usepackage{xspace}
\usepackage{comment}

\sloppy

\def\transarrow{\xrightarrow}
\newcommand{\setarrow}[1]{\def\transarrow{#1}}

\newcommand{\trule}[2]{\frac{#1}{#2}}
\newcommand{\crule}[3]{\frac{#1}{#2},\;{#3}}
\newcommand{\withenv}[2]{{#1}\vdash{#2}}
\newcommand{\trans}[3]{{#1}\transarrow{#2}{#3}}
\newcommand{\ctrans}[4]{{#1}\transarrow{#2}{#3},\;{#4}}
\newcommand{\llang}[1]{\mbox{\lstinline[mathescape]|#1|}}
\newcommand{\pair}[2]{\inbr{{#1}\mid{#2}}}
\newcommand{\inbr}[1]{\left<{#1}\right>}
\newcommand{\highlight}[1]{\color{red}{#1}}
\newcommand{\ruleno}[1]{\eqno[\scriptsize\textsc{#1}]}
\newcommand{\inmath}[1]{\mbox{$#1$}}
\newcommand{\lfp}[1]{fix_{#1}}
\newcommand{\gfp}[1]{Fix_{#1}}
\newcommand{\vsep}{\vspace{-2mm}}
\newcommand{\supp}[1]{\scriptsize{#1}}
\newcommand{\G}{\mathfrak G}
\newcommand{\sembr}[1]{\llbracket{#1}\rrbracket}
\newcommand{\cd}[1]{\texttt{#1}}
\newcommand{\miniKanren}{\texttt{miniKanren}\xspace}
\newcommand{\ocanren}{\texttt{ocanren}\xspace}

\lstdefinelanguage{ocanren}{
keywords={fresh, let, in, match, with, when, class, type,
object, method, of, rec, repeat, until, while, not, do, done, as, val, inherit,
new, module, sig, deriving, datatype, struct, if, then, else, open, private, virtual, include, success, failure},
sensitive=true,
commentstyle=\small\itshape\ttfamily,
keywordstyle=\ttfamily\underbar,
identifierstyle=\ttfamily,
basewidth={0.5em,0.5em},
columns=fixed,
fontadjust=true,
literate={fun}{{$\lambda$}}1 {->}{{$\to$}}3 {===}{{$\equiv$}}1 {=/=}{{$\not\equiv$}}1 {|>}{{$\triangleright$}}3 {/\\}{{$\wedge$}}2 {\\/}{{$\vee$}}2 {^}{{$\uparrow$}}1,
morecomment=[s]{(*}{*)}
}

\lstset{
mathescape=true,
basicstyle=\small,
identifierstyle=\ttfamily,
keywordstyle=\bfseries,
commentstyle=\scriptsize\rmfamily,
basewidth={0.5em,0.5em},
fontadjust=true,
language=ocanren
}

\usepackage{letltxmacro}
\newcommand*{\SavedLstInline}{}
\LetLtxMacro\SavedLstInline\lstinline
\DeclareRobustCommand*{\lstinline}{%
  \ifmmode
    \let\SavedBGroup\bgroup
    \def\bgroup{%
      \let\bgroup\SavedBGroup
      \hbox\bgroup
    }%
  \fi
  \SavedLstInline
}

\begin{document}

\mainmatter

\title{MiniKanren Semantics for Programs Transformation}

\maketitle

\section{Syntax}

\begin{figure}
\centering 
{\bf Supplementary syntax categories:}
$$
\begin{array}{rcll}
  \mathcal{X} &=& x_1,\, x_2,\dots & \mbox{\supp(syntactic variables)} \\
  \mathcal{S} &=& \sigma_1,\, \sigma_2,\dots & \mbox{\supp(semantic variables)} \\
  \mathcal{R}^n &=& r^n_1, r^n_2, \dots &\mbox{\supp{(predicate names indexed by the arity n)}}
\end{array}
$$

{\bf Terms:}
$$
\begin{array}{rcll}
  \mathcal{T(V)} &=& \mathcal{V} \,|\, C^n \, t_1,\dots,t_n & \mbox{\supp(terms parameterized by variable type)} \\
  \mathcal{T_X} &=& \mathcal{T(X)} & \mbox{\supp(terms which contain only syntactic variables)} \\
  \mathcal{T_S} &=& \mathcal{T(S)} & \mbox{\supp(terms which contain only syntactic variables)} 
\end{array}
$$

{\bf Goals:}
$$
\begin{array}{rcll}
  \mathcal{G} &=& t_1 \equiv t_2 & \mbox{\supp(unification)} \\
              &|& g_1 \wedge g_2 & \mbox{\supp(conjunction)} \\
              &|& g_1 \vee   g_2 & \mbox{\supp(disjunction)} \\
              &|&\lstinline|fresh $x$ $\;g$| & \mbox{\supp{(fresh logical variable binder)}}\\
              &|& \mathcal{R}^{n} \, t_1 ,\dots, t_n & \mbox{\supp{(relations)}} \\
\end{array}
$$

{\bf Definitions:}
$$
\begin{array}{rcll}
  \mathcal{D}^n &=& \lambda x_1, \dots, x_n. g & \mbox{\supp(relation definition)}
\end{array}
$$

{\bf States:}
$$
\begin{array}{rcll}
  I &=& \mathcal{X} \rightarrow \mathcal{T_S} & \mbox{\supp(syntax variables interpretation)} \\
  \Sigma &=& \mathcal{S} \rightarrow \mathcal{T_S} & \mbox{\supp(substitutions)} \\
  \Delta &=& 2^\mathcal{S} & \mbox{\supp(the set of used semantic variables)} \\
  \mathfrak{S} &=& I \times \Sigma \times \Delta & \mbox{\supp(states)} \\
  \Gamma &=& \mathcal{R}^n \rightarrow \mathcal{D}^n & \\
  \Gamma &\vdash& \mathfrak{s} \xrightarrow{g} \mathfrak{\bar{s}} & \mbox{\supp(semantics of the goal g)}
\end{array} 
$$

\end{figure}

\newpage
\section{Semantics}

\begin{figure}
\centering 
$$
\withenv{\Gamma}{ \ctrans{(\iota,\sigma,\delta)}{t_1 \equiv t_2}{[]}{mgu \, (\iota \, t_1) \sigma \,\, (\iota \, t_2) \sigma = \bot} } 
$$

$$
\withenv{\Gamma}{ \trans{(\iota,\sigma,\delta)}{t_1 \equiv t_2}{(\iota, mgu \, (\iota \, t_1) \sigma \; (\iota \, t_2) \sigma , \delta)}}	
$$

$$
\crule{\withenv{\Gamma}{\trans{(\iota', \sigma, \delta')}{g}{\mathfrak{\bar{s}}}}}
      {\withenv{\Gamma}{\trans{(\iota, \sigma, \delta)}{\lstinline|fresh $\;x$ $\;g$|}{\mathfrak{\bar{s}}}}}
      {\iota' = \iota[x \leftarrow s], \delta' = \delta \cup \{x\}, x \notin \delta}
$$

$$
\crule{\withenv{\Gamma}{\trans{(\iota',\sigma,\delta)}{g}{\mathfrak{\bar{s}}}}}
      {\withenv{\Gamma}{\trans{(\iota,\sigma,\delta)}{r^n \, t_1 \dots t_n}{\mathfrak{\bar{s}}}}}
      {\iota' = \iota[x_i \leftarrow t_i \iota], \Gamma \, r^n = \lambda x_1 \dots x_n. g}
$$

$$
\begin{array}{rcll}
  \square \iota &:& \mathcal{T_X} \rightarrow \mathcal{T_S} & \\
  x \iota &=& \iota (x) & \\
  \mathcal{C}^n \, t_1 \dots t_n &=& \mathcal{C}^n \, (t_1 \iota) \dots (t_n \iota)
\end{array} 
$$

$$
\trule{\withenv{\Gamma}{\trans{\mathfrak{s}}{g_1}{\mathfrak{\bar{s_1}}}} \;\;\;\; \withenv{\Gamma}{\trans{\mathfrak{s}}{g_2}{\mathfrak{\bar{s_2}}}}}
      {\withenv{\Gamma}{\trans{\mathfrak{s}}{g_1 \vee g_2}{\mathfrak{\bar{s_1}} \oplus \mathfrak{\bar{s_2}}}}}
$$

$$
\trule{\withenv{\Gamma}{\trans{(\iota,\sigma,\delta)}{g_1}{\mathfrak{\bar{s_1}}}} \;\;\;\; \withenv{\Gamma}{\trans{\mathfrak{\bar{s_1}}}{g_2 \iota}{\mathfrak{\bar{s_2}}}}}
      {\withenv{\Gamma}{\trans{(\iota,\sigma,\delta)}{g_1 \wedge g_2}{\mathfrak{\bar{s_2}}}}}
$$

$$
\trule{\withenv{\Gamma}{\trans{\mathfrak{s}}{g}{\mathfrak{\bar{s_1}}}} \;\;\;\; \withenv{\Gamma}{\trans{\mathfrak{\bar{s}}}{g}{\mathfrak{\bar{s_2}}}}}
      {\withenv{\Gamma}{\trans{\mathfrak{s}:\mathfrak{\bar{s}}}{g}{\mathfrak{\bar{s_1}} \oplus \mathfrak{\bar{s_2}}}}}
$$

\end{figure}

\end{document}