### 2020-05-21 papers/miniKanren/text/conclusion.tex
Draft for miniKanren workshop

\section{Conclusion}

We compared some of the specialization techniques for \mk{}.

There seem to be no one good technique.