### 2020-05-21 papers/miniKanren/text/intro.tex

\section{Introduction}

The \mk{} language family is nice.

Unpredictable running time for different directions.

Specialization sometimes helps, sometimes does not, sometimes makes everything worse.

Control issues, blah blah blah

This problem also appear in supercompilation (partial evaluation, specialization) of functional and imperative programming languages.

<Here should be review of problems in specialization of functional languages>

We came up with the new specialization algorithm, which seems nice, but is as unpredictable as the others.

Here are some numbers...

Contribution is the following
\begin{itemize}
  \item explanation of specialization
  \item new specialization algorithm
  \item why some programs behave worse after specialization
\end{itemize}