### 2020-05-21 papers/miniKanren/text/intro.tex
about transformers disagv

\section{Introduction}

The \mk{} language family is nice.

Unpredictable running time for different directions.

Specialization sometimes helps, sometimes does not, sometimes makes everything worse.

Control issues, blah blah blah

This problem also appear in supercompilation (partial evaluation, specialization) of functional and imperative programming languages.

While the aim of all meta-computation techniques is residual program efficiency, they rarely consider the residual program efficiency from the target language evaluator point of view.
The main optimization source is precomputing all possible intermediate and statically-known semantics steps at transformation-time while other criteria like residual program size or possible optimizations and execution cost of different language constructions by target language evaluator are usually out of consideration~\cite{jonesbook}.
It is known that supercompilation may adversely affect GHC optimizations yielding standalone compilation more powerful~\cite{SCBE,TCES} and cause code explosion~\cite{SCHC}.
Moreover, caused by transformer complexity it may be hard to predict real program speedup in each case on a bunch of examples even disregarding problems above.
Wherein, for example, for partial evaluation, an apogee of useless is the use of all static variables in a dynamic context only while it is known to be usefull for some specific form of language processors~\cite{jonesbook,bulyonkov84}.
Anyway, the formalization problem of cases for useful transformers application is poorly studied.

<Here should be review of problems in specialization of functional languages>

We came up with the new specialization algorithm, which seems nice, but is as unpredictable as the others.

Here are some numbers...

Contribution is the following
\begin{itemize}
  \item explanation of specialization
  \item new specialization algorithm
  \item why some programs behave worse after specialization
\end{itemize}