### 2020-05-21 papers/miniKanren/text/evaluation.tex
miniKanren workshop draft

\section{Evaluation}

In the evaluation we compared the performance of our CPD, Ecce and the new non-conjunctive partial evaluator.
We have also implemented the branching heuristic instead of the deterministic one in the CPD, just to be sure, it is not a cureall.

We used the following 4 programs to test our specializers on.
\begin{itemize}
  \item propositional evaluator
  \begin{itemize}
    \item Compicated bool expression in the last conjunctive
    \item Table implementation of the boolean operation, placed as a first conjunct
  \end{itemize}
  \item searching for a unifier of two terms
  \item searching for paths of a specific length in a graph
\end{itemize}

The first one demonstrates the drawback of the approach implemented in ECCE: left-to-right deterministic unfolding.
The boolean expression which narrows the first conjunts is the last, so ECCE does not unfold it until too late.

The second one should be easier for this approach.

The unification is a difficult problem, specialization of which is of importance.