### 2020-05-22 papers/miniKanren/text/evaluation.tex
WIP

\section{Evaluation}

In our study we compared the performance of our implementation of CPD, ECCE and the new non-conjunctive partial evaluator.
We have also implemented the branching heuristic instead of the deterministic one in the CPD, just to ensure, it is not a cureall.

We used the following 4 programs to test the specializers on.
\begin{itemize}
  \item Two implementations of an evaluator of logic formulas.
  \item A program to compute a unifier of two terms.
  \item A program to search for paths of a specific length in a graph
\end{itemize}

All these relations are relational interpreters.
Their last argument is a boolean value which is $true$ if the other arguments are in relation and $false$ otherwise.

\subsection{Evaluator of Logic Formulas}

The relation $eval^o$ describes evaluation of a subset of first-order logic formulas in a given substitution.
$eval^o$ has 3 arguments.
The first argument is a list of boolean values which serves as a substitution.
The $i$-th value of the list is the value of the $i$-th variable.

The second argument is a formula with the following abstract syntax.
A formula is either a \emph{variable} represented with a Peano number, a \emph{negation} of a formula, a \emph{conjunction} of two formulas or a \emph{disjunction} of two formulas.

The third argument is a boolean value which is $true$ if the given formula in the given substition is true and $false$ otherwise.

One possible implementation of the evaluator in the syntax of \oc{} is presented in listing~\ref{eval:last}.
Here the relation \lstinline{elem$^o subst v res$} unifies \lstinline{res} with the value of the variable \lstinline{v} in the list \lstinline{subst}.
The relations \lstinline{and$^o$}, \lstinline{or$^o$}, and \lstinline{not$^o$} encode corresponding boolean operations.

\begin{lstlisting}[label={eval:last}, caption={Evaluator of formulas with boolean operation last}, captionpos=b]
  let rec eval$^o$_last subst fm res = conde [
    fresh (x y z v w) (
      (fm === var v /\ elem$^o$ subst v res);
      (fm === conj x y /\ eval$^o$_last st x v /\ eval$^o$_last st y w /\ and$^o$ v w res);
      (fm === disj x y /\ eval$^o$_last st x v /\ eval$^o$_last st y w /\ or$^o$ v w res);
      (fm === neg x /\ eval$^o$_last st x v /\ not$^o$ v res))]
\end{lstlisting}

Note, that the calls to boolean relations \lstinline{and$^o$}, \lstinline{or$^o$}, and \lstinline{not$^o$} are placed last within each conjunction.
This poses a challenge to the CPD-based specializers.
Conjunctive partial deduction unfolds relation calls from left to right, so when specializing this relation for running backwards (i.e. considering the goal \lstinline{eval$^o$ subst fm ^true}), it fails to propagate the direction data onto recursive calls of \lstinline{eval$^o$}.
It leads to over-unfolding, big residual programs and less than optimal performance.

Moving boolean operations to the left, as shown in listing~\ref{eval:fst}, we get a program which is easier for CPD to specialize.

\begin{lstlisting}[label={eval:fst}, caption={Evaluator of formulas with boolean operation second}, captionpos=b]
  let rec eval$^o$_plain subst fm res = conde [
    fresh (x y z v w) (
      (fm === var v /\ elem$^o$ subst v res);
      (fm === conj x y /\ and$^o$_table v w res /\ eval$^o$_plain st x v /\ eval$^o$_plain st y w);
      (fm === disj x y /\ or$^o$_table v w res /\ eval$^o$_plain st x v /\ eval$^o$_plain st y w);
      (fm === neg x /\ not$^o$_table v res /\ eval$^o$_plain st x v))]
  \end{lstlisting}

The next complication for partial deduction is how much sequential unfoldings should be performed to  restrict other calls in a conjunction in a meaningful way.
In this example the way in which boolean relations are implemented affect the specialization result.
The simplest way to implement these relations is with a table as shown in listing~\ref{not:table}

\begin{lstlisting}[label={not:table}, caption={Implementation of boolean \lstinline{not} as a table}, captionpos=b]
  let not$^o$_table x y = conde [
     (x === ^true /\ y === ^false;
      x === ^false /\ y === ^true)]
\end{lstlisting}

The other way to implement boolean operations is via one base boolean relation such as \lstinline{nand$^o$} which is in turn has a table-based implementation.

\begin{lstlisting}[label={not:table}, caption={Implementation of boolean operation via \lstinline{nand}}, captionpos=b]
  let not$^o$ x y = nand$^o$ x x y

  let or$^o$ x y z = nand$^o$ x x xx /\ nand$^o$ y y yy /\ nand$^o$ xx yy z

  let and$^o$ x y z = nand$^o$ x y xy /\ nand$^o$ xy xy z

  let nand$^o$ a b c = conde [
    ( a === ^false /\ b === ^false /\ c === ^true );
    ( a === ^false /\ b === ^true /\ c === ^true );
    ( a === ^true /\ b === ^false /\ c === ^true );
    ( a === ^true /\ b === ^true /\ c === ^false )]
\end{lstlisting}

We considered two implementations of \lstinline{eval$^o$}: \lstinline{eval$^o$_plain} and \lstinline{eval$^o$_last} and studied how specializers behave on them.
The first one uses table-based boolean operations and places them at the second place in each conjunction.
The relation \lstinline{eval$^o$_last} employs boolean operations implemented via \lstinline{nand$^o$} and place them at the end of each conjunction.
These two programs are complete opposites from the standpoint of CPD which the time measurements in table~\ref{tbl:eval} confirmed.

We measured time necessary to generate $10000$ formulas over two variables which evaluate to \lstinline{^true}.
We compared the result of specialization of the goal \lstinline{eval$^o$ subst fm ^true} by our implementation of CPD, done by ECCE system and the new non-conjunctive partial deduction.
Since ECCE cannot work directly on \mk{} programs, we first translated them to \pro{} and then back: this translation is purely syntactical.

\begin{table}
  \centering
  \begin{tabular}{c|c|c}
                   & last & plain \\
  \hline\hline
  Original         & >60.00s & >60.00s \\
  \hline
  Ecce             & 7.12s  & 9.22s \\
  \hline
  CPD              & 31.31s & 5.46s \\
  \hline
  Non-CPD          & 4.99s  & 5.05s \\
  \hline
  Branches         & 17.21s  & 6.17s \\
  \hline
  \end{tabular}

  \caption{Searching for 10000 true logic formulas}
  \label{tbl:eval}
\end{table}

\subsection{Search for a Unifier}

\begin{table}
  \centering
  \begin{tabular}{c|c}

    & running time \\
  \hline\hline
  Original         & >300.00s \\
  \hline
  CPD              & 2.35s \\
  \hline
  Non-CPD          & 14.90s \\
  \hline
  Branches         & >300.00s \\
  \hline
  \end{tabular}

  \caption{Searching for a unifier for terms f(X, X, g(Z, t)) and f(g(p, L), Y, Y)}
  \label{tbl:unify}
\end{table}

The unification of two terms $t$ and $u$ is searching for a substitution $\theta$ such that $t \theta = u \theta$, $\theta$ is called a unifier.
We search for any unifier, not necessarily most specific.
The details on this benchmark can be found in the earlier paper CITE.
This example demonstrates how too much unfolding can be introduced with the non-conjunctive partial deduction.

\subsection{Search for Paths in a Graph}

Here we search for 5 paths in a graph.
The details on this benchmark can be found in the earlier paper CITE.

\begin{table}
  \centering
  \begin{tabular}{c|c}

    & running time \\
  \hline\hline
  Original         & 19.86s \\
  \hline
  CPD              & 4.66s \\
  \hline
  Non-CPD          & 3.00s \\
  \hline
  \end{tabular}

  \caption{Searching for paths in a graph}
  \label{tbl:unify}
\end{table}