### 2020-05-22 papers/miniKanren/text/abstract.tex
Introduction

\begin{abstract}
  We explore partial deduction for \mk{}: a specialization technique aimed at improving the performance of a relation in the given direction.
  We describe a novel approach to specialization of \mk{} based on partial deduction and supercompilation.
  On several examples, we demonstrate issues which arise during partial deduction.
\end{abstract}