### 2020-05-22 papers/miniKanren/text/conclusion.tex
Evaluation and conclusion

\section{Conclusion}

In this paper we discussed some issues which arise in partial deduction of a relational programming language \mk{}.
We presented a novel approach to partial deduction which uses a heuristic to select the most suitable relation call to unfold at each step of driving.
We compared this approach to the earlier implementation of conjunctive partial deduction.
Our evaluation showed that there is still not one good technique which definitively speeds up every relational program.