### 2020-07-06 papers/aplas/text/conclusion.tex
Conclusion rewritten

\section{Conclusion}

In this paper we discussed some issues which arise in the area of partial deduction techniques for relational programming language \mk.
We presented a novel approach to partial deduction --- conservative partial deduction --- which uses a heuristics to select a suitable relation call to unfold at each step of driving.
We compared this approach with the most sophisticated implementation of conjunctive partial deduction --- \ecce partial deduction system --- on 10 relations which solve 6 different problems.

Our specializer improved the execution time of 10 queries to the  relations, while worsening 2 queries and not affecting one query.
\ecce worsened the performance of all 4 implementations of the propositional evaluator relation, while improving other queries.
Conservative PD is more stable with regrads to the order of relation calls than \ecce which is demonstrated by the similar performance of all 4 implementations of the evaluator of logic formulas.

Some queries to the same relation were improved better by ConsPD, while others --- by \ecce.
We conclude that there is still no one good technique which definitively speeds up every relational program.
More research is needed to develop models capable of predicting the performance of a relation which can be used in specialization.