### 2020-07-06 papers/aplas/text/figures/heuristic.tex

\documentclass[12pt,a4paper]{report}
\usepackage[lmargin=3.81cm,tmargin=2.54cm,rmargin=2.54cm,bmargin=2.52cm]{geometry}
\usepackage{multirow}
\usepackage{color,xcolor}
\definecolor{darkblue}{rgb}{0,0,.75}
\usepackage{listings} %iclude code in your document
\definecolor{dkgreen}{rgb}{0,0.6,0}

\lstloadlanguages{Matlab} %use listings with Matlab for Pseudocode
\lstnewenvironment{PseudoCode}[1][]
{\lstset{basicstyle=\scriptsize,
    keywordstyle=\bfseries\underbar,
    literate=
      {:=}{$\leftarrow{}$}{1}
      {==}{$={}$}{1},
    numbers=left,
    morekeywords={ifd,define,where,let,in,error,case,of,while,do,otherwise,
      read,goto,return,build-cons,build-cst,build-if,if,then,else,do,switch,of},
    morestring=[b]",
    morestring=[s][]{\#\{}{\}},
    deletestring=[b]{"},
    alsoletter={-},
    frame=l,
    keywordstyle = [2]{\textit},
    morekeywords = [2]{lookup,initial_code,extend,reduce,eval,rest,first_command,first_instruction,new_tail,first_symbol},
    showspaces=false,
    showstringspaces=false,
    stringstyle=\color{blue},
    identifierstyle=,
    escapechar=@*,
    xleftmargin=.04\textwidth,#1}}
{}

\usepackage{stmaryrd}
\begin{document}

\thispagestyle{empty}
\begin{figure}[h]
\begin{PseudoCode}[moredelim={[is][stringstyle]{@@}{@@}},
  moredelim={[is][\color{dkgreen}\bfseries]{@!}{@!}},
  moredelim={[is][\color{red}\bfseries]{!!}{!!}}]
heuristically_select_a_call :: Conjunction @*$\to$@* Maybe Call
heuristically_select_a_call C = find heuristics C

heuristics :: Call @*$\to$@* Bool
heuristics r = isStatic r || isDeterministic r || isLessBranching r
\end{PseudoCode}
\end{figure}

\end{document}