### 2020-07-07 papers/aplas/text/abstract.tex
Abstract

\begin{abstract}
  We study conjunctive partial deduction, an advanced specialization technique aimed at improving the performance of logical programs, in the context of relational programming language \mk. We identify a number of issues, caused by  \mk pecularities, and describe a novel approach to specialization based on partial deduction and supercompilation. While the results of the evaluation do not yet demonstrate the advantages of our approach in all cases, we nevertheless consider it as the first step towards an efficient optimization framework for \mk.
\end{abstract}