### 2020-07-17 papers/miniKanren/text/abstract.tex
Camera-ready miniKanren'20

\begin{abstract}
  We explore partial deduction, an advanced specialization technique aimed at improving the performance of a relation in the given direction, in the context of \mk{}.
  On several examples, we demonstrate issues which arise during partial deduction of relational programs.
  We describe a novel approach to specialization of \mk{} based on partial deduction and supercompilation.
  Although the proposed approach does not give the best results in all cases, we view it as a stepping stone towards the efficient optimization of \mk{}.

\end{abstract}