### 2020-08-17 papers/miniKanren/slides/doubleAppPD.tex
Slides: many figures

\begin{tikzpicture}[
  processTree,
  level 3/.style={sibling distance=23em},
  level 4/.style={sibling distance=10em}]
  \node {\rel{doubleAppend}{$x \ y \ z \ r$}}
    child { node {\rel{append}{$x \ y \ t$} \and \rel{append}{$t \ z \ r$}}
      child { node[diamond] {\and}
        child { node (1) {\rel{append}{$x \ y \ t$} }
          child { node {\subst{$x \to [], y \to t$}} }
          child { node (12) {\rel{appendo}{$x' \ y \ t'$} \\ \subst{$x \to h :: x', t \to h :: t'$}}}}
        child { node (2) {\rel{appendo}{$t \ z \ r$} }
          child { node (21) {\subst{$t \to [], z \to r$}} }
          child { node (22) {\rel{appendo}{$t' \ y \ r'$} \\ \subst{$t \to h :: t', r \to h :: r'$}}}}}};
  \draw [dashed,<-] (1.east) to [out=10,in=30] (12.east);
  \draw [dashed,<-] (2.east) to [out=10,in=30] (22.east);
  \draw [red,<->] (12.south) to [out=300,in=240] (21.south);
\end{tikzpicture}