### 2020-08-17 papers/miniKanren/slides/propCPDunf3.tex

\begin{tikzpicture}[
  processTree,
  sibling distance=10em,
  level distance=8em]
  \node {
    \underline{\rel{eval}{$fm \ s \ true$}}}
    child { node[draw=none, fill=none] {...}}
    child { node {
      \rel{eval}{$x \ s \ a$} \and \\
      \underline{\rel{not}{$a \ true$}} \\
      \subst{$fm \to neg \ x$}}
      child { node {
        \rel{eval}{$x \ s \ false$} \\
        \subst{$a \to false$}}}}
    child { node[draw=none, fill=none] {...}}
    child { node[draw=none, fill=none] {...}}
  ;
\end{tikzpicture}