### 2020-08-19 papers/miniKanren/slides/doubleAppPD.tex
More slides

\begin{tikzpicture}[
  processTree,
  level 3/.style={sibling distance=23em},
  level 4/.style={sibling distance=10em},
  anchor=center]
  \node
    {\rel{double\_append}{$x \ y \ z \ r$}}
    child { node {\rel{append}{$x \ y \ t$} \and \rel{append}{$t \ z \ r$}}
      child { node (d) [diamond] {\and}
        child { node (1) {\rel{append}{$x \ y \ t$} }
          child { node {\subst{$x \to [], y \to t$}} }
          child { node (12) {\rel{appendo}{$x' \ y \ t'$} \\ \subst{$x \to h :: x', t \to h :: t'$}}}}
        child { node (2) {\rel{appendo}{$t \ z \ r$} }
          child { node (21) {\subst{$t \to [], z \to r$}} }
          child { node (22) {\rel{appendo}{$t' \ y \ r'$} \\ \subst{$t \to h :: t', r \to h :: r'$}}}}}};

  \draw [densely dotted,->] ($(d.east)+(5,1.5)$) node[above,draw=none] {\Large split node} to [out=210,in=0] (d.east);

  \draw [dashed,<-] (1.east) to [out=10,in=30] (12.east);
  \draw [dashed,<-] (2.east) to [out=10,in=30] (22.east);

  \draw [red,<->] ($(12.south)+(0.7,0.15)$) to [out=300,in=240] node[below, draw=none, black] {\Large substitutions conflict} ($(21.south)+(-0.65,0.15)$);

\end{tikzpicture}