### 2020-08-25 papers/miniKanren/slides/evalUnfold.tex

\begin{tikzpicture}[
  processTree,
  anchor=center]
  \node (root) {\rel{eval}{$fm \ s \ true$}}
    child { node {$fm \equiv conj \ x \ y $ \and \rel{and}{$a \ b \ true$} \and \rel{eval}{$x \ s \ a$} \and \rel{eval}{$y \ s \ b$}}
      child { node {
        \rel{and}{$a \ b \ true$} \and
        \rel{eval}{$x \ s \ a$} \and
        \rel{eval}{$y \ s \ b$} \\
        \subst{$fm \to conj \ x \ y$}}
    }
    }
    ;

  \node[left=4em of root, yshift=-0.5cm] (lookup) {...};
  \draw [->] (root.west) to [out=-170,in=10] (lookup.east);

  \node[right=4em of root, yshift=-0.5cm] (lookup) {...};
  \draw [->] (root.east) to [in=170,out=-10] (lookup.west);
\end{tikzpicture}