### 2020-08-25 papers/miniKanren/slides/propCPDNoSplit.tex

\begin{tikzpicture}[
  processTree,
  anchor=center,
  level 2/.style={level distance=5em},
  level 3/.style={level distance=4em},
  sibling distance=10em]
  \node (root) {\rel{eval}{$fm \ s \ true$}}
      child { node {
        \underline{\rel{and}{$a \ b \ true$}} \and
        \rel{eval}{$x \ s \ a$} \and
        \rel{eval}{$y \ s \ b$} \\
        \subst{$fm \to conj \ x \ y$}}
        child { node {
          \underline{\rel{eval}{$x \ s \ true$}} \and
          \rel{eval}{$y \ s \ true$} \\
          \subst{$a \to true, b \to true$}}
          child {node {
            \underline{\rel{and}{$x' \ y' \ true$}} \and
            \rel{eval}{$x' \ s \ a'$} \and
            \rel{eval}{$y' \ s \ b'$} \and
            \rel{eval}{$y \ s \ b$} \\
            \subst{$x \to conj \ x' \ y'$}
          }
          child { node (grow) {
            \underline{\rel{eval}{$x' \ s \ a'$}} \and
            \rel{eval}{$y' \ s \ b'$} \and
            \rel{eval}{$y \ s \ b$} \\
            \subst{$a' \to true, b' \to true$}}
            child {node[draw=none,fill=none] {...}}
          }
        }
      }}
    ;

  \node[left=4em of root, yshift=-0.5cm] (lookup) {...};
  \draw [->] (root.west) to [out=-170,in=10] (lookup.east);

  \node[right=4em of root, yshift=-0.5cm] (lookup) {...};
  \draw [->] (root.east) to [in=170,out=-10] (lookup.west);

  \draw [red,->] ($(grow.east)+(1,-2.5)$) node[below,draw=none,anchor=north east] {\Large uncontrollable growth} to [out=45,in=0] (grow.east);

\end{tikzpicture}