### 2020-08-25 papers/miniKanren/slides/propCPDSplit.tex

\begin{tikzpicture}[
  processTree,
  anchor=center,
  level 2/.style={level distance=5em},
  level 3/.style={level distance=4em},
  level 4/.style={level distance=3em},
  sibling distance=10em]
  \node (root) {\rel{eval}{$fm \ s \ true$}}
      child { node {
        \underline{\rel{and}{$a \ b \ true$}} \and
        \rel{eval}{$x \ s \ a$} \and
        \rel{eval}{$y \ s \ b$} \\
        \subst{$fm \to conj \ x \ y$}}
        child { node {
          \rel{eval}{$x \ s \ true$} \and
          \rel{eval}{$y \ s \ true$} \\
          \subst{$a \to true, b \to true$}}
          child { node [diamond] {\and}
            child { node (x) {
              \rel{eval}{$x \ s \ true$}}
            }
            child { node (y) {
              \rel{eval}{$y \ s \ true$}}}
          }
        }
      }
    ;

  \node[left=4em of root, yshift=-0.5cm] (lookup) {...};
  \draw [->] (root.west) to [out=-170,in=10] (lookup.east);

  \node[right=4em of root, yshift=-0.5cm] (lookup) {...};
  \draw [->] (root.east) to [in=170,out=-10] (lookup.west);

  \draw[dashed,->] (x.west) to [out=165,in=-165] (root.west);
  \draw[dashed,->] (y.east) to [out=15,in=-15] (root.east);

\end{tikzpicture}