### 2020-08-25 papers/miniKanren/slides/propSimpl.tex

\begin{tikzpicture}[
  processTree,
  sibling distance=7em,
  level distance=7em,
  level 2/.style={level distance=5em},
  anchor=center]
  \node {
    \rel{eval}{$fm \ s \ true$}}
    child { node[draw=none, fill=none] {...}}
    child { node {
      \rel{and}{$a \ b \ true$} \and  \\
      \rel{eval}{$x \ s \ a$} \and \\
      \rel{eval}{$y \ s \ b$} \\
      \subst{$fm \to conj \ x$}}
      child { node[draw=none, fill=none] {...}}
      }
    child { node[draw=none, fill=none] {...}}
  ;
\end{tikzpicture}